---
title: "Lab2"
author: "Eliott Doyle, Diana DeWald"
date: "1/24/2022"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fivethirtyeight)
library(fivethirtyeightdata)

# View(google_trends)

```

## 1. Start with the google trends data

### >> Visualize the change in trends for each hurricane over time in one plot with three scales. Map color to a discrete scale.

```{r q1.1, echo = FALSE, warning = FALSE, message = FALSE}

# plotting change in trends over time with three scales
google_trends %>%
  pivot_longer(c("hurricane_harvey_us", "hurricane_irma_us", "hurricane_maria_us", "hurricane_jose_us"),
               names_to = "Hurricane", 
               values_to = "Searches") %>%
  ggplot(aes(date, Searches)) +
  geom_line(aes(color = Hurricane), se = FALSE, size = 2) +
  theme_classic() +
  labs(title = "U.S. Google trends for Hurricanes", subtitle = "Aug.-Sep. 2017", x = "Date", y = "Search Frequency (thousands)")+
  scale_color_manual(labels = c("Harvey", "Irma", "Maria", "Jose"),  values = c("darkcyan", "plum4", "lightsalmon", "darkseagreen4")) +
  guides(color=guide_legend("Hurricane"))

```

#### The three scales for our plot are: google search frequency on y-axis (we assume that the values in the dataset were thousands), date on the x-axis, and hurricane on a color scale.

### >> Visualize the trends again, this time with solid shapes (if your first plot was already solid shapes, use lines). Hint, geom_area can be difficult to work with if you have a grouping factor. Try geom_ribbon instead.

```{r q1.2, echo = FALSE, warning = FALSE, message = FALSE}
google_trends$Harvey <- google_trends$hurricane_harvey_us
google_trends$Irma <- google_trends$hurricane_irma_us
google_trends$Jose <- google_trends$hurricane_jose_us
google_trends$Maria <- google_trends$hurricane_maria_us

google_trends %>%
  pivot_longer(c("Harvey", "Irma", "Maria", "Jose"),
               names_to = "Hurricane", 
               values_to = "Searches") %>%
  ggplot(aes(date, Searches)) +
  geom_point(aes(shape = Hurricane), alpha = 0.4, se = FALSE, size = 2) +
  theme_classic() +
  labs(title = "U.S. Google trends for Hurricanes", subtitle = "Aug.-Sep. 2017", x = "Date", y = "Search Frequency (thousands)")

```


### >> Visualize the same relation, but change the scales so color is mapped to a continuous scale (the other scales should be discrete).

```{r q1.3, echo = FALSE, warning = FALSE, message = FALSE}
library(viridis)

google_trends$Harvey <- google_trends$hurricane_harvey_us
google_trends$Irma <- google_trends$hurricane_irma_us
google_trends$Jose <- google_trends$hurricane_jose_us
google_trends$Maria <- google_trends$hurricane_maria_us

google_trends %>%
  pivot_longer(c("Harvey", "Irma", "Maria", "Jose"),
               names_to = "Hurricane", 
               values_to = "Searches") %>%
  ggplot(aes(date, Hurricane)) +
  geom_tile(aes(fill = Searches, stat="identity", width=1, height=.9)) + 
  theme_classic() +
  labs(title = "U.S. Google trends for Hurricanes", subtitle = "Aug.-Sep. 2017", x = "Date", y = "Hurricane") +
  scale_fill_viridis()

```

### >> Create a data frame (using tibble() or data.frame()) that includes the dates that each hurricane made landfall. Annotate one of your first two plots to include these reference points with vertical lines (Hint: use geom_vline(), specifying the data source for this layer as this new dataset). You do not have to include text annotations - just the vertical lines, but if you would like to try, look into geom_text() or geom_label(). The hurricanes made landfall on (approximately) the following days:
* Harvey: August 25, 2017
* Irma: September 10, 2017
* Jose: Did not make landfall
* Maria: September 20, 2017

```{r q1.4, echo = FALSE, warning = FALSE, message = FALSE}
# creating the dataframe
Hurricane <- c("Harvey", "Irma", "Jose", "Maria")
date <- c("2017-08-25", "2017-09-10", "NA", "2017-09-20")
landfall_dates <- data_frame(Hurricane, date)
landfall_dates$date <- as.Date(landfall_dates$date)
#View(landfall_dates)

# modifying the line graph from q1.1
google_trends %>%
  rename(harvey = hurricane_harvey_us,
         irma = hurricane_irma_us,
         maria = hurricane_maria_us, 
         jose = hurricane_jose_us) %>% 
  pivot_longer(c("Harvey", "Irma", "Maria", "Jose"),
               names_to = "Hurricane", 
               values_to = "Searches") %>%
  ggplot(aes(date, Searches)) +
  geom_line(aes(color = Hurricane), se = FALSE, size = 2) +
  theme_classic() +
  labs(title = "U.S. Google trends for Hurricanes", subtitle = "Aug.-Sep. 2017", x = "Date", y = "Search Frequency (thousands)")+
  scale_color_manual(values = c("darkcyan", "plum4", "lightsalmon", "darkseagreen4")) +
  guides(color=guide_legend("Hurricane")) +
  geom_vline(data = landfall_dates, aes(xintercept = date))
```
### Make any final modifications to the plot you’d like to communicate the information as effectively as possible

```{r q1.5, echo = FALSE, warning = FALSE, message = FALSE}
google_trends %>%
  rename(harvey = hurricane_harvey_us,
         irma = hurricane_irma_us,
         maria = hurricane_maria_us, 
         jose = hurricane_jose_us) %>% 
  pivot_longer(c("Harvey", "Irma", "Maria", "Jose"),
               names_to = "Hurricane", 
               values_to = "Searches") %>%
  ggplot(aes(date, Searches)) +
  geom_line(aes(color = Hurricane), se = FALSE, size = 2) +
  theme_classic() +
  labs(title = "U.S. Google trends for Hurricanes", subtitle = "Aug.-Sep. 2017", x = "Date", y = "Search Frequency (thousands)")+
  scale_color_manual(values = c("darkcyan", "plum4", "lightsalmon", "darkseagreen4")) +
  guides(color=guide_legend("Hurricane")) +
  geom_vline(data = landfall_dates, aes(xintercept = date, color = Hurricane), size = 1.3)

```

